---
title: "Capstone-EFA"
author: "Annaliese Bronner"
date: "10/26/2020"
output: html_document
---

```{r}
#install.packages("psych")
library(psych)

#install.packages("tidyverse")
library(tidyverse)

library(dplyr)

```

```{r}
setwd("C:/Users/annal.000/Desktop/Microbiology Capstone/Project Data")
ARS_data<- read.csv("ARS.csv")
Socio_data <- read.csv("Socioeconomic Status.csv")
Covid_data <- read.csv("Covid_data.csv")
Excess_Deaths <- read.csv("WyoExcessDeathsbyCounty.csv")

names(Socio_data)[names(Socio_data) == "State...County.Name"] <- "County"
names(Socio_data)[names(Socio_data) == "All.Ages.in.Poverty.Percent"] <- "Percent_in_Poverty"
Recent_Socio_data<-filter(Socio_data, Year == "2018")

Sex_data <- data.frame(ARS_data$County, ARS_data$Percent.Female, ARS_data$Percent.Male)

Combined_Sex_data <- read.csv("ARS_CombinedSexData.csv")

Selected_Socio_data <- select(Recent_Socio_data, Year, County, Percent_in_Poverty)

Combined_Race_data <- read.csv("ARS_CombinedRaceData.csv")

SEM_Trial_1_data <- data.frame(ARS_data, Selected_Socio_data$Percent_in_Poverty)

names(SEM_Trial_1_data)[names(SEM_Trial_1_data) == "Percent_in_Poverty"] <- "Percent.in.Poverty"

Nonwhite_data<-read.csv("Race-Ethnicity-Non White.csv")

AQI<- read.csv("Jared_Data.csv")

SEM_Trial_2_data <- data.frame(ARS_data, Selected_Socio_data$Percent_in_Poverty, Covid_data$Cases.Per.Hundred.Thousand,Covid_data$Deaths,Nonwhite_data$Non.White, AQI$Overall.AQI, Excess_Deaths$October)

names(SEM_Trial_2_data)[names(SEM_Trial_2_data) == "Covid_data.Cases.Per.Hundred.Thousand"] <- "Cases.Per.Hundred.Thousand"
names(SEM_Trial_2_data)[names(SEM_Trial_2_data) == "Selected_Socio_data.Percent_in_Poverty"] <- "Percent.in.Poverty"
names(SEM_Trial_2_data)[names(SEM_Trial_2_data) == "Covid_data.Deaths"]<-"Deaths"
names(SEM_Trial_2_data)[names(SEM_Trial_2_data) == "Nonwhite_data.Non.White"] <- "Non.White"                        

No_Albany<-SEM_Trial_2_data[SEM_Trial_2_data$County !="Albany",]

No_Outliers <- SEM_Trial_2_data[SEM_Trial_2_data$County !="Carbon" & SEM_Trial_2_data$County !="Campbell",]

Median_Age_Covid_Plot<-read.csv("Age and Covid Bar Graph.csv")  


write.csv(SEM_Trial_2_data, "SEM_Final_Dataset.csv")
```


```{r}
library(ggplot2)

ggplot(data=ARS_data, aes(County,Median.Age)) +
  geom_point() +
  theme_minimal() +
  xlab("County") +
  ylab("Median Age")+
   theme(axis.text.x = element_text(size = 10, angle = 60, hjust = 1)) 

ggplot(data=ARS_data, aes(County,Population.Density)) +
  geom_point() +
  theme_minimal() +
  xlab("County") +
  ylab("Population Density")+
   theme(axis.text.x = element_text(size = 10, angle = 60, hjust = 1))


#install.packages("ggthemes")
#install.packages("extrafont")
#install.packages("plyr")
#install.packages("scales")
library(ggplot2)
library(ggthemes)
library(extrafont)
library(plyr)
library(scales)


ggplot(Combined_Sex_data, aes(x = County, y = Percent, fill = Sex)) + 
  geom_bar(stat = "identity") +
  xlab("\nCounty") +
  ylab("Percent of Population\n") +
  guides(fill = FALSE) +
  theme(axis.text.x = element_text(size = 10, angle = 60, hjust = 1)) +
  geom_text(data=Combined_Sex_data, aes(x = County, y = Percent,label = paste0(Percent,"%")), size=2)


ggplot(Combined_Race_data, aes(x = County, y = Percent, fill = Race),show.legend=TRUE) +
  geom_bar(stat = "identity") +
  theme(legend.position="right",legend.direction="vertical") + 
  xlab("\nCounty") +
  ylab("Percent of Population\n") +
  guides(fill = FALSE) +
  theme(axis.text.x = element_text(size = 10, angle = 60, hjust = 1))+
  theme(legend.text=element_text(size=14),legend.title=element_text(size=14))

```

This code was a first trial to get a hang of the lavaan package, before I had all of the data needed for our actual models. The single squiggle means a one way analysis, the double squiggle means a two way analysis. Example: Median.Age with respect to American.Indian AND American.Indian with respect to Median.Age.

This code can be disregarded. 
```{r}
#install.packages("lavaan")
#install.packages("semPlot")
#install.packages("dplyr")


library(lavaan)  #for doing the CFA
library(semPlot)  #for plotting your CFA
library(dplyr)  #for subsetting data quickly if needed

summary(m2<-lm(Percent.in.Poverty ~ Median.Age + American.Indian.and.Alaska.Native.alone, data=SEM_Trial_2_data))

m2  <- 'Percent.in.Poverty ~ Median.Age + American.Indian.and.Alaska.Native.alone
        Median.Age ~~ American.Indian.and.Alaska.Native.alone'
m2.fit <- sem(m2, data=SEM_Trial_2_data)
summary(m2.fit, stand=TRUE, rsq=TRUE)
semPaths(m2.fit, what="std", whatLabels="std", residuals=FALSE, laBout="circle2")

m3 <- 'Percent.in.Poverty ~ Median.Age + American.Indian.and.Alaska.Native.alone
        Median.Age ~ American.Indian.and.Alaska.Native.alone'
m3.fit <- sem(m3, data=SEM_Trial_2_data)
summary(m3.fit, stand=TRUE, rsq=TRUE)
semPaths(m3.fit, what="std", whatLabels="std", residuals=FALSE, laBout="circle2")
modificationindices(m3.fit)

```


This is the code for the actual SEM models. Here I've run several different models to later run an AIC backwards selection to determine the best fit model. 
```{r}

summary(m4<-lm(Cases.Per.Hundred.Thousand ~ Median.Age + Percent.in.Poverty, data=SEM_Trial_2_data))


m4  <- 'Cases.Per.Hundred.Thousand ~ Median.Age + Percent.in.Poverty
        Median.Age ~~ Percent.in.Poverty'
m4.fit <- sem(m4, data=SEM_Trial_2_data)
summary(m4.fit, stand=TRUE, rsq=TRUE)
semPaths(m4.fit, what="std", whatLabels="std", residuals=FALSE, laBout="circle2")



ggplot(SEM_Trial_2_data, aes(x=Median.Age, y=Cases.Per.Hundred.Thousand)) +
  geom_bar(stat="identity") +
  xlab("Median Age")+
  ylab("Cases Per 100,000")

#In order to have both covid cases and covid deaths on the same graph I had to run a log transformation on both numbers. This made the difference in sizes significantly smaller and created a better visual. This took several tries, way more than it should have, but you can almost always find what you need on Google. The most difficult thing was figuring out how my data set needed to be formatted. 
ggplot(Median_Age_Covid_Plot, aes(Median.Age, log(People),fill=Number)) +
  geom_bar(stat="identity", position = "dodge")+
  xlab("Median Age Group")+
  ylab("Individuals")

 ggplot(No_Albany, aes(Median.Age, Cases.Per.Hundred.Thousand))+
   geom_point()
 
 summary(m6<-lm(Cases.Per.Hundred.Thousand~ Non.White + Percent.in.Poverty + Median.Age + Population.Density, data=SEM_Trial_2_data))


m6 <- 'Cases.Per.Hundred.Thousand ~ Non.White + Percent.in.Poverty + Median.Age + Population.Density
        Non.White ~~ Percent.in.Poverty
        Median.Age ~~ Percent.in.Poverty
        Median.Age ~~ Non.White
        Population.Density ~~ Non.White
        Population.Density ~~ Percent.in.Poverty
        Population.Density ~~ Median.Age'
m6.fit <- sem(m6, data=SEM_Trial_2_data)
summary(m6.fit, stand=TRUE, rsq=TRUE)
semPaths(m6.fit, what="std", whatLabels="std", residuals=FALSE)
modificationindices(m6.fit)

m7 <- 'Cases.Per.Hundred.Thousand~Non.White+Median.Age+AQI.Overall.AQI
        Non.White ~~ Median.Age
        Median.Age ~~ AQI.Overall.AQI'
m7.fit <- sem(m7, data=SEM_Trial_2_data)
summary(m7.fit, stand=TRUE, rsq=TRUE)
semPaths(m7.fit, what="std", whatLabels="std", residuals=FALSE)
modificationindices(m7.fit)
```


```{r}
#AIC Backwards Selection for model 5

library(readr)
library(lme4)

summary(m5<-lm(Cases.Per.Hundred.Thousand~ Non.White + Percent.in.Poverty + Median.Age + Population.Density + Non.White*Percent.in.Poverty + Non.White*Median.Age + Population.Density*Median.Age + AQI.Overall.AQI, data=SEM_Trial_2_data))


m5 <- 'Cases.Per.Hundred.Thousand ~ Non.White + Percent.in.Poverty + Median.Age + Population.Density + AQI.Overall.AQI
        Non.White ~~ Percent.in.Poverty
        Median.Age ~~ Percent.in.Poverty
        Median.Age ~~ Non.White
        Population.Density ~~ Non.White
        Population.Density ~~ Percent.in.Poverty
        Population.Density ~~ Median.Age
        Non.White ~~ AQI.Overall.AQI
        Median.Age ~~ AQI.Overall.AQI
        Population.Density ~~ AQI.Overall.AQI
        Percent.in.Poverty ~~ AQI.Overall.AQI'
m5.fit <- sem(m5, data=SEM_Trial_2_data)
summary(m5.fit, stand=TRUE, rsq=TRUE)
semPaths(m5.fit, what="std", whatLabels="std", residuals=FALSE)
modificationindices(m5.fit)

m5<-lm(Cases.Per.Hundred.Thousand~ Non.White + Percent.in.Poverty + Median.Age + Population.Density + Non.White*Percent.in.Poverty + Non.White*Median.Age + Population.Density*Median.Age + AQI.Overall.AQI, data=SEM_Trial_2_data)


#install.packages("MuMIn")
library(MuMIn)
AIC(m5)  #AIC Rate= 295.1616


Reduced_5.1 <- lm(Cases.Per.Hundred.Thousand ~ Non.White + Percent.in.Poverty + Median.Age +Population.Density + AQI.Overall.AQI +  Non.White*Percent.in.Poverty + Non.White*Median.Age , data=SEM_Trial_2_data) #AIC Rate = 294.0964	

Reduced_5.2 <- lm(Cases.Per.Hundred.Thousand ~ Non.White + Percent.in.Poverty + Median.Age + Population.Density + Non.White*Median.Age + Population.Density*Median.Age + AQI.Overall.AQI, data=SEM_Trial_2_data) #AIC Rate = 293.2799	

Reduced_5.3 <- lm(Cases.Per.Hundred.Thousand ~ Non.White + Percent.in.Poverty +Median.Age + Population.Density + Non.White*Percent.in.Poverty + Population.Density*Median.Age + AQI.Overall.AQI, data=SEM_Trial_2_data)  #AIC Rate = 293.6813		
AIC(Reduced_5.1, Reduced_5.2, Reduced_5.3)

#The full model is the best fit for covid rate, so we will run no more reductions.


Reduced_5.4 <- lm(Cases.Per.Hundred.Thousand~ Non.White + Percent.in.Poverty + Median.Age + Population.Density + AQI.Overall.AQI, data=SEM_Trial_2_data)
AIC(Reduced_5.4) #AIC Rate = 291.209

Reduced_5.5 <- lm(Cases.Per.Hundred.Thousand ~ Non.White + Percent.in.Poverty + Median.Age + AQI.Overall.AQI, data=SEM_Trial_2_data)
AIC(Reduced_5.5) #AIC Rate = 290.0484

Reduced_5.6 <- lm(Cases.Per.Hundred.Thousand ~ Non.White + Percent.in.Poverty + AQI.Overall.AQI, data=SEM_Trial_2_data)
AIC(Reduced_5.6) #AIC Rate = 288.828

Reduced_5.7 <- lm(Cases.Per.Hundred.Thousand ~ Median.Age + Percent.in.Poverty + AQI.Overall.AQI, data=SEM_Trial_2_data)
AIC(Reduced_5.7) #AIC Rate = 289.2202

Reduced_5.8 <- lm(Cases.Per.Hundred.Thousand~ Non.White +  Median.Age, data=SEM_Trial_2_data)
AIC(Reduced_5.8) #AIC Rate = 367.8357

Reduced_5.9 <- lm(Cases.Per.Hundred.Thousand~ Non.White + Median.Age + AQI.Overall.AQI, data=SEM_Trial_2_data)
AIC(Reduced_5.9) #AIC Rate = 290.9076

Reduced_5.10 <- lm(Cases.Per.Hundred.Thousand~ Non.White +  Median.Age + Population.Density, data=SEM_Trial_2_data)
AIC(Reduced_5.10) #AIC Rate = 369.0629

#Reduced 5.6 has the lowest AIC, so we will accept this as our best model for COVID rate. This only considers non-white percent of population and median age.
```

```{r}


summary(m8<-lm(Deaths~ Non.White + Percent.in.Poverty + Median.Age + Population.Density + Non.White*Percent.in.Poverty + Non.White*Median.Age + Population.Density*Median.Age + AQI.Overall.AQI, data=SEM_Trial_2_data))


m8 <- 'Deaths~ Non.White + Percent.in.Poverty + Median.Age + Population.Density +AQI.Overall.AQI
        Non.White ~~ Percent.in.Poverty
        Median.Age ~~ Percent.in.Poverty
        Median.Age ~~ Non.White
        Population.Density ~~ Non.White
        Population.Density ~~ Percent.in.Poverty
        Population.Density ~~ Median.Age
        Non.White ~~ AQI.Overall.AQI
        Median.Age ~~ AQI.Overall.AQI
        Population.Density ~~ AQI.Overall.AQI
        Percent.in.Poverty ~~ AQI.Overall.AQI'
m8.fit <- sem(m8, data=SEM_Trial_2_data)
summary(m8.fit, stand=TRUE, rsq=TRUE)
semPaths(m8.fit, what="std", whatLabels="std", residuals=FALSE)
modificationindices(m8.fit)


m8<-lm(Deaths~ Non.White + Percent.in.Poverty + Median.Age + Population.Density + Non.White*Percent.in.Poverty + Non.White*Median.Age + Population.Density*Median.Age + AQI.Overall.AQI, data=SEM_Trial_2_data)

#install.packages("MuMIn")
library(MuMIn)
AIC(m8) #AIC Deaths= 101.0684


Reduced_8.1 <- lm(Deaths ~ Non.White + Percent.in.Poverty + Median.Age +Population.Density + Non.White*Percent.in.Poverty + Non.White*Median.Age +AQI.Overall.AQI, data=SEM_Trial_2_data)#AIC Deaths = 102.6117 

Reduced_8.2 <- lm(Deaths ~ Non.White + Percent.in.Poverty + Median.Age + Population.Density + Non.White*Median.Age + Population.Density*Median.Age +AQI.Overall.AQI, data=SEM_Trial_2_data) #AIC = 103.4470	 	

Reduced_8.3 <- lm(Deaths ~ Non.White + Percent.in.Poverty +Median.Age + Population.Density + Non.White*Percent.in.Poverty + Population.Density*Median.Age +AQI.Overall.AQI, data=SEM_Trial_2_data) #AIC = 102.9221 
AIC(Reduced_8.1, Reduced_8.2, Reduced_8.3)

#The full model is the best fit for covid deaths, so we will run no more reductions.

```


```{r}
summary(m9<-lm(Excess_Deaths.October~ Non.White + Percent.in.Poverty + Median.Age + Population.Density + Non.White*Percent.in.Poverty + Non.White*Median.Age + Population.Density*Median.Age + AQI.Overall.AQI, data=SEM_Trial_2_data))


m9 <- 'Excess_Deaths.October~ Non.White + Percent.in.Poverty + Median.Age + Population.Density +AQI.Overall.AQI
        Non.White ~~ Percent.in.Poverty
        Median.Age ~~ Percent.in.Poverty
        Median.Age ~~ Non.White
        Population.Density ~~ Non.White
        Population.Density ~~ Percent.in.Poverty
        Population.Density ~~ Median.Age
        Non.White ~~ AQI.Overall.AQI
        Median.Age ~~ AQI.Overall.AQI
        Population.Density ~~ AQI.Overall.AQI
        Percent.in.Poverty ~~ AQI.Overall.AQI'
m9.fit <- sem(m9, data=SEM_Trial_2_data)
summary(m9.fit, stand=TRUE, rsq=TRUE)
semPaths(m9.fit, what="std", whatLabels="std", residuals=FALSE)
modificationindices(m9.fit)


```


```{r}
summary(m10<-lm(Deaths~ Non.White + Percent.in.Poverty + Median.Age + Population.Density + Non.White*Percent.in.Poverty + Non.White*Median.Age + Population.Density*Median.Age + AQI.Overall.AQI, data=No_Outliers))


m10 <- 'Deaths~ Non.White + Percent.in.Poverty + Median.Age + Population.Density +AQI.Overall.AQI
        Non.White ~~ Percent.in.Poverty
        Median.Age ~~ Percent.in.Poverty
        Median.Age ~~ Non.White
        Population.Density ~~ Non.White
        Population.Density ~~ Percent.in.Poverty
        Population.Density ~~ Median.Age
        Non.White ~~ AQI.Overall.AQI
        Median.Age ~~ AQI.Overall.AQI
        Population.Density ~~ AQI.Overall.AQI
        Percent.in.Poverty ~~ AQI.Overall.AQI'
m10.fit <- sem(m10, data=No_Outliers)
summary(m10.fit, stand=TRUE, rsq=TRUE)
semPaths(m10.fit, what="std", whatLabels="std", residuals=FALSE)
modificationindices(m10.fit)

```

```{r}
library(rsconnect)
rsconnect::deployApp('Capstone-EFA.Rmd')
summary(m11<-lm(Cases.Per.Hundred.Thousand~ Non.White + Percent.in.Poverty + Median.Age + Population.Density + Non.White*Percent.in.Poverty + Non.White*Median.Age + Population.Density*Median.Age + AQI.Overall.AQI, data=No_Outliers))


m11 <- 'Cases.Per.Hundred.Thousand ~ Non.White + Percent.in.Poverty + Median.Age + Population.Density + AQI.Overall.AQI
        Non.White ~~ Percent.in.Poverty
        Median.Age ~~ Percent.in.Poverty
        Median.Age ~~ Non.White
        Population.Density ~~ Non.White
        Population.Density ~~ Percent.in.Poverty
        Population.Density ~~ Median.Age
        Non.White ~~ AQI.Overall.AQI
        Median.Age ~~ AQI.Overall.AQI
        Population.Density ~~ AQI.Overall.AQI
        Percent.in.Poverty ~~ AQI.Overall.AQI'
m11.fit <- sem(m11, data=No_Outliers)
summary(m11.fit, stand=TRUE, rsq=TRUE)
semPaths(m11.fit, what="std", whatLabels="std", residuals=FALSE)
modificationindices(m11.fit)

```


Calculating the correlation coefficient for covid cases per 100,000 and different factors. There are currently no strong correlations, some moderate. The correlation between covid cases and age changes quite a lot when taking out Albany county, which has the lowest median age and highest covid rate.(11/3/2020)
```{r}
 cor(No_Albany$Median.Age, No_Albany$Cases.Per.Hundred.Thousand)
 
 cor(SEM_Trial_2_data$Population.Density,SEM_Trial_2_data$Cases.Per.Hundred.Thousand)
 
 cor(SEM_Trial_2_data$Cases.Per.Hundred.Thousand, SEM_Trial_2_data$Percent.in.Poverty)
 
 cor(SEM_Trial_2_data$Cases.Per.Hundred.Thousand, SEM_Trial_2_data$AQI.Overall.AQI, na.rm=TRUE)
 
 cor(SEM_Trial_2_data$Cases.Per.Hundred.Thousand, SEM_Trial_2_data$American.Indian.and.Alaska.Native.alone)
 
  cor(SEM_Trial_2_data$Cases.Per.Hundred.Thousand, SEM_Trial_2_data$White.alone)
```


This is how you can export a data frame you've made and convert it into a csv 11/17/20
```{r}
write.csv(SEM_Trial_2_data, "SEM_Trial_2_data.csv")
```

